\documentclass{article}

\renewcommand*\familydefault{\sfdefault} 


\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm, txfonts, stmaryrd}
\usepackage{bussproofs}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{corollary}[theorem]{Corollary}

\title{On the Consistency of Height and Width Potentialism}
\author{Chris Scambler}
\date{\today}
\begin{document}
\maketitle
\abstract{
    Recent work in philosophy of set theory has furnished 
    some arguments that height and width potentialism 
    are inconsistent with one another. 
    One such argument can be found in this volume (Brauer); 
    the other is forthcoming (Roberts).

	At the same time, others have suggested 
    there may be some merit in the combination of 
    height  and width  potentialism. Such authors 
    have presented views that appear to manifest 
    the combination in a non-trivial way, and have 
    defended philosophical claims on their basis 
    (e.g. that all sets are ultimately countable 
    -- cf Builes \& Wilson, Meadows, 
    Scambler, Pruss).  

	Clearly there is a tension here. 
    The business of this article is to explain 
    (what I take to be) its solution. 
    I will argue that height  and width  potentialism 
    are compatible, and that there is  
    even an attractive view in the foundations of 
    mathematics that arises from their combination. 
    I will do this by explaining that view and how it 
    responds to the arguments alleging inconsistency.

}
\section{Overview}
Recent work in philosophy of set theory has furnished 
some arguments that height  and width  potentialism 
are inconsistent with one another. 
One such argument can be found in this volume (Brauer); 
the other is forthcoming (Roberts).

At the same time, others have suggested 
there may be some merit in the combination of 
height  and width  potentialism. Such authors 
have presented views that appear to manifest 
the combination in a non-trivial way, and have 
defended philosophical claims on their basis 
(e.g. that all sets are ultimately countable 
-- cf Builes \& Wilson, Meadows, 
Scambler, Pruss).  

Clearly there is a tension here. 
The business of this article is to explain 
(what I take to be) its solution. 
I will argue that height  and width  potentialism 
are compatible, and that there is  
even an attractive view in the foundations of 
mathematics that arises from their combination. 
I will do this by explaining that view and how it 
responds to the arguments alleging inconsistency.

The plan is as follows. 
Section 1 gives some definitional background and context. 
Section 2 then presents some axiomatic systems that purport to 
combine height  and width  potentialism and discusses
issues of relevant to consistency, citing new results
on strength and equivalence to standard `actualist' 
set theories. Section 3 presents the 
inconsistency argument of Roberts and explains 
how a proponent of height and width potentialism should reply.
Section 4 does the same for Brauer.

\section{Background}
Let me start be explaining the terms. 

For present purposes, \emph{potentialism} in set theory 
will be the idea that there could always be 
more sets than there in fact are. 
In slogan form, for the potentialist,  
the universe of sets is `indefinitely extendible'. 

\emph{Height potentialism} is the idea that the universe
is always extendible `upwards' to include 
new sets of higher rank than any given ones. 
\emph{Width potentialism}, on the other hand, 
is the idea that the universe is always extendible 
`outwards', to contain new sets of no greater rank 
than the max of any given ones.

Historically, height potentialism has been motivated 
by considerations involving the set-theoretic paradoxes. 
Russell's paradox shows there is no set of all 
non-self-membered sets, and hence that the cumulative 
hierarchy of all sets is itself (therefore) not a set. 
But the height potentialist complains that any 
`stopping point' for the cumulative hierarchy 
would be arbitrary. Surely there is no conceptual obstacle 
to any particular collection of ranks of the cumulative 
hierarchy providing `urelements' for a longer continuation.

It is natural to
explicate this idea of an indefinitely extendible 
universe of sets in modal logic. 
Indeed, modal axiom systems based around core height potentialist 
ideas are known to exhibit tight forms of 
equivalence with standard iterative set theories like ZFC.

Width potentialism has been historically 
less popular, although has been attracting 
some attention over the last few decades as a 
way understand the independence phenomenon in set 
theory. There is, in fact, a close structural 
similarity in the motivation for height  and 
width  potentialism in these terms (cf Meadows). 
Just as the height potentialist begins with the 
intuition that it is arbitrary that there should be 
some ranks of the cumulative hierarchy that somehow 
inherently can't extended to include further sets, 
the width  potentialist may begin with the idea that 
it is arbitrary that there should be some universe of 
sets that cannot be extended by forcing over its 
partial orders. Just as in the previous case, the 
mathematics of forcing seems to lead us to believe 
there is at least no conceptual obstacle to making 
sense of such `forcing extensions of the universe'.

Again, as with height potentialism, a natural way for the 
width  potentialist to formalize their view 
involves modal logic: one formulates an axiom to the 
fact that, for any partial order $\mathbb{P}$, it is possible to 
find a generic for $\mathbb{P}$. Such explicitly 
axiomatic approaches 
to width  potentialism are not very well-studied: 
the focus of most work in this area has been on model 
theory. Nevertheless, axiomatic approaches are possible 
and easy enough to formulate. 

There are, in any event, clear analogies 
in the cases for height  and width  potentialism. 
In each case one has a central inexistence result 
in first order set theory (Russell's paradox, the 
proof that some filters do not admit partial orders) 
and one seeks to overcome it, after a fashion, by 
implementing a modalized version -- any things can 
form a set in the first case, any partial order can 
be forced over in the second. Indeed, it is natural to think 
that going potentialist one way might give you some 
reason to consider going potentialist the other way too. 

\section{Height and Width Potentialism Combined}

In this section, I will present some axiom systems that seem to 
explicate the intuitive idea of height and width potentialism, as 
described above.

\subsection{Core Logical Principles}
All the versions of height and width potentialism we will consider will be built around
two principles: the first is the height potentialist principle that any things can be 
the elements of a set; the second is the width potentialist `forcing axiom', that 
a generic filter can be found for any partial order. 

Both potentialist principles involve modality; the first also involves plural 
quantification. Accordingly the language $\mathcal{L}_0$
we use to formulate these theories 
will have at least a modal operator $\Box$, singular variables $x_n$, plural variables 
$X_n$, the propositional connectives and quantifiers, the identity symbol $=$ 
and the symbol $\in$ for set membership. The circumstance that some thing $x$ is 
one of some things $X$ will be represented by the concatenation $Xx$. Identity 
is only well formed between terms of the same type (singular/plural). 

The following core axioms will be included in all the potentialist theories we 
will go on to discuss.
\begin{description}
    \item[FQ] Standard rules for free quantifier logic 
    with identity for each type of variable.
    \item[Mod] The modal logic $\mathsf{S4.2}$ with converse Barcan formula,
    necessitation, and standard rules for universal quantification within the scope of $\Box$.  
    \item[P-ext] $\Box \forall x [\Diamond Xx \equiv \Diamond Yx] \supset X = Y$
    \item[PR] The plural rigidity axioms 
    \begin{enumerate} 
        \item $ \Diamond Xx \rightarrow \Box Xx$ 
        \item $ \Diamond (\exists x Xx \wedge x = y) \rightarrow \exists x Xx \wedge x = y$.
    \end{enumerate}
    \item[Comp] For any formula $\varphi(x, y, Y)$, 
    $\Box \forall y \Box \forall Y \Box \exists X \forall x [Xx \leftrightarrow \varphi(x, y, Y)]$ is an axiom.
    \item[P-Choice] A plural version of the axiom of choice.
    \item[M-Rep] A modal version of the axiom scheme of replacement.
    \item[Inf] An axiom saying that the set of natural numbers exists.
    \item[Set-Ext] The axiom of extensionality in the form 
    $\Box \forall x [\Diamond x \in y \equiv \Diamond x \in z] \supset y = z $
    \item[Sets] The axiom of foundation, and an axiom 
      $\Box \forall x \exists X \Box \forall y[Xy \leftrightarrow y \in x]$ asserting the rigidity of set membership
\end{description}
The precise details of many of these won't be too relevant below, but they are cited 
here for something like completeness' sake. More details can be found in CITE. The 
odd one out, perhaps, is {\bf Inf}. It is included here for simplicity, since 
many of the 
phenomena we will be interested in occur only at the level of infinite sets.

The basic 
idea behind the core axioms is to get a modal and plural logic, combined with fundamental 
principles of set theory, to allow for the development of set theory based on potentialist 
principles of set existence. Let us now turn to that development.

\subsection{Simple HWP}
In this section I will present what seems to me to be the simplest formal combination 
of height and width potentialism. The theory is not mathematically strong, being 
exactly equivalent in strength to second order arithmetic. But it does offer a clean 
and simple proof of concept for the combination of height and width potentialism, 
as well as a useful starting point for further extensions.

The first axiom we add is:
\begin{description}
    \item[HP] $\Box \forall X \Diamond \exists x \forall y[y \in x \leftrightarrow Xy]$
    \begin{itemize} 
    \item \emph{Any things can be the elements of a set.}
    \end{itemize}
\end{description}
HP enshrines the core idea of height potentialism, since it implies (given the
usual notion of rank) that given any sets one can find others of still higher rank.

The second requires a little more preparation. In our background logic we 
can define the notion of a partial order, the notion of being a filter on 
a partial order, and the notion of being a dense set in a partial order in 
the standard way. Let blackboard variables $\mathbb{P}, \mathbb{Q}$ range over 
partial orders. (These are singular variables.) Let $D(X, \mathbb{P})$ mean 
that all $X$s are dense in $\mathbb{P}$. Finally, let $F_\cap(x, \mathbb{P}, X)$ 
mean that $x$ is a filter on $\mathbb{P}$ that intersects every one of the $X$s.
Then our width potentialist axiom can be stated as:
\begin{description}
    \item[WP] $\Box \forall \mathbb{P} \forall X D(X, \mathbb{P}) \rightarrow \Diamond \exists g(F_\cap(g, \mathbb{P}, X)) $   \begin{itemize} 
    \item \emph{Any partial order can be used in forcing.}
    \end{itemize}
\end{description}
WP is one way to flesh out width potentialism, since it implies that any 
sufficiently rich pluarlity of sets
sets may be extended to include ones of rank no greater than their max by 
forcing. 

The following will be useful to us going forward. In it, $f : \mathbb{N} \twoheadrightarrow X$
is an abbreviation for the assertion that $f$ is a function on $\mathbb{N}$ with every $X$ in its range.

\begin{proposition}\label{SHWP0}
    Let Count  be the principle:
    $$\Box \forall X \Diamond \exists f[f : \mathbb{N} \twoheadrightarrow X] $$ 
    then Count is equivalent to WP over the core logic + HP.
\end{proposition}
\begin{proof}
    See CATBC.
\end{proof}
Let us call the result of adding these HP and either WP or Count
 to the core logic 
 Simple Height and Width Potentialism, or SHWP.

Turning now to questions of consistency, 
it turns out that SHWP is 
demonstrably consistent relative to Second Order Arithmetic (SOA). 
In fact it turns out that SHWP exhibits a tight form of equivalence with SOA,
something Tim Button has called `near synonymy' in the 
recent literature. 

(Here and below, we understand SOA under the guise of 
ZFC without power + all sets are countable. This is 
definitionally equivalent to the more standard
arithmetical formulations; see e.g. Simpson.)

A detailed account of the relationship between SOA and SHWP would be overkill here.
(Details can be found in CITE.) But certain features will be important 
to our discussion of consistency below, and having some idea of how the interpretation 
of the modal theory SHWP in the non-modal SOA goes will be very 
helpful to us going forward. 
So we will spend some time discussing certain aspects of the `tight equivalence'
just mentioned.

Let $\mathcal{L}_\in$ be the first order 
language of set theory. Then:
\begin{theorem}\label{SHWP1}
    There is a map $\cdot^\Diamond : \mathcal{L}_\in \to \mathcal{L}_0$ that preserves 
    theoremhood from SOA to SHWP.
\end{theorem}
\begin{theorem}\label{SHWP2}
    There is a map $\cdot^\exists : \mathcal{L}_0 \to \mathcal{L}_\in$ that preserves 
    theoremhood from SHWP to
    SOA.
\end{theorem}
Theorem \ref{SHWP1} says that SHWP interprets SOA; this 
direction does not really concern us directly, since it does not 
imply anything about the consistency of SHWP, but some 
idea of how things go will be useful.
The translation $\varphi \mapsto \varphi^\Diamond$ proceeds by prepending every 
universal quantifier in $\varphi$ by a $\Box$ and every existential by a $\Diamond$.
A result due to Linnebo says that for $\varphi \in \mathcal{L}_\in$, 
\[\Gamma \vdash \varphi \Leftrightarrow \Gamma^\Diamond \vdash_{Core} \varphi^\Diamond\]
where $\Gamma^\Diamond$ has the obvious meaning of $\{\gamma^\Diamond : \gamma \in \Gamma\}$.
It thus suffices to prove $\varphi^\Diamond$ for each axiom of SOA in SHWP, 
which (using Proposition \ref{SHWP0}) is not difficult. 

Theorem \ref{SHWP2} is directly on-topic. It says that SOA interprets SHWP and hence
secures the consistency of the latter relative to the former. Full details of the 
ideal proof are a little fiddly and would go beyond our needs here. But some idea, 
again, of how things go down will be useful.

The key idea behind the translation $\varphi \mapsto \varphi^\exists$ is to factor 
out use of modal operators in favor of quantification over possible worlds. 

In a bit more detail, the idea is that we define the notion of a possible world 
in SOA to be a transitive set. That means we interpret the language of SHWP so that ordinary 
quantifiers are always restricted to such worlds, and the modal operators induce 
quantifiers over such worlds. 

Formally, this means our translation will have to 
carry formulas $\varphi$ in the modal language (which may be sentences) to formulas 
$\varphi^\exists(w)$ 
in the first order language with a free `world' variable $w$. Key clauses of the 
translation are things like
\[ (\forall x \varphi)^\exists(w) := \forall x \in w \varphi^\exists(w) \]
\[ (\forall X \varphi)^\exists(w) := \forall x \subseteq w \varphi^\exists(w) \]
\[ (\Box \varphi)^\exists(w) := \forall u [Tran(u) \wedge w \subseteq u \rightarrow \varphi^\exists(u)]\]
where $Tran(u)$ of course is the assertion that $u$ is transitive. (The case of 
plural containment $Xx$ is a little fiddly, but can be made to work quite nicely.)
The potentialist axioms are then readily seen to be true, under the translation, 
in SOA. For HP, this is because every set is an element of a transitive set. 
For WP, this is because SOA proves all sets are countable. So for any 
given partial order, we may move to a transitive set that witnesses its countability, 
and then proceed to introduce a generic for it if need be.

A final point is that these results can be strengthened to attain something 
approximating definitional equivalence between the two theories.  ...
\subsection{Strong HWP}
The theory I have just cited combines height and width potentialism in a 
straightforward way. The result is a theory that is (up to a near-synonymy) 
second order arithmetic. This is a pretty weak theory, although as we know 
from the reverse mathematics literature it is plenty strong enough to develop 
much of the mathematics needed in applications. Can the height and with potentialist 
do better?

There is indeed a reasonable way to proceed here. Our potentialist is interested 
with expansions of the universe along two `directions'. One has the ability 
to extend the universe `upwards' to create sets of higher rank; and one has the ability 
to extend it `outwards' by forcing. By separating out these two possible methods of 
expansion, and by asserting strong axioms regarding what can be attained along the 
vertical dimension of expansion alone, much stronger HWP systems can be derived,
indeed ones with all the power of ZFC and more. This extension makes the general 
picture of height and width potentialism much more philosophically interesting, since 
such versions have at least the consistency strength to recover all standard mathematics.
But they also bring new conceptual problems with them, problems that will be exploited 
in one of the inconsistency arguments we will consider. 

Let us first discuss in a little more detail how to implement this idea. We expand 
the language $\mathcal{L}_0$ to $\mathcal{L}_1$ by adding in two new modal operators, 
$\circledbar$ and $\circledbslash$. $\circledbar$ is to reflect possibility by \emph{only} vertical expansion: 
one can think of this as possibility by \emph{only} iteratively introducing new sets 
for given pluarlities. $\circledbslash$, on the other hand, is to reflect possibility by \emph{only}
horizontal expansion: one can think of this, indifferently, as just by adding 
new generic filters, or adding new enumerating functions. ($\boxbar$ and $\boxbslash$
are the duals.) $\Diamond$ remains in the 
language, and represents `absolute' possibility, that is, possibility by either 
domain expansion method. 

Here is a way to axiomatize a `strong' system of this kind.\footnote{I ignore 
fiddly details with the background multi-modal logic. Details can be found in CITE.}

We begin by modifying HP and WP. 
\begin{description}
    \item[HP] $\Box \forall X \circledbar \exists x \forall y[y \in x \leftrightarrow Xy]$
\end{description}

\begin{description}
    \item[WP] $\Box \forall \mathbb{P} \forall X D(X, \mathbb{P}) \rightarrow \circledbslash \exists g(F_\cap(g, \mathbb{P}, X)) $
\end{description}

Next, we add rules to reflect the generality of $\Diamond$.
\begin{description}
    \item[Gh] $\circledbar \varphi \rightarrow \Diamond \varphi$
    \item[Gw] $\circledbslash \varphi \rightarrow \Diamond \varphi$
\end{description}
Finally, add the following restricted version of the 
powerset axiom to vertical possibility.
\begin{description}
    \item[r-Pow] $\Box \forall x \circledbar \exists y \boxbar \forall z[ z \in y \leftrightarrow z \subseteq x]$
\end{description}
r-Pow says that it is always possible by vertical expansion to get all the subsets of 
any given set, \emph{so long as one disregards the subsets you can get by forcing}. 
This is a kind of local powerset axiom: in intuitive (procedural) terms,
it says that you can always eventually get 
all the subsets of a given set you can get without forcing, if you go on introducing 
sets long enough.

It is useful to compare r-Pow with the corresponding `unrestricted' version 
\begin{description}
    \item[Pow] $\Box \forall x \circledbar \exists y \Box \forall z[ z \in y \leftrightarrow z \subseteq x]$
\end{description}
which is provably inconsistent with {\bf WP} over the rest of the theory. Pow says that 
it is possible for there to such things as \emph{all possible} subsets of any given 
set; but given the existence of an infinite set (as we are guaranteed), and the 
universal possibility of forcing, enshrined in {\bf WP}, this cannot be, since we can always force 
to add new subsets to any given set.

Let the extension of the core theory by the above principles (but not, of course, Pow)
be called HWP. The following facts come easily.
\begin{theorem}\label{HWP0}
    There is a translation $\cdot^\circledbar : \mathcal{L}_\in \to \mathcal{L}_1$
    that preserves theoremhood from ZFC to HWP.
\end{theorem}
\begin{theorem}\label{HWP1}
    There is a translation $\cdot^\Diamond : \mathcal{L}_\in \to \mathcal{L}_1$
    that preserves theoremhood from SOA to HWP.
\end{theorem}
In each case, the result follows much as it did in the previous case. In fact 
Theorem \ref{HWP1} really just is Theorem \ref{SHWP1}. The first uses the same translation
but with $\circledbar$ in place of $\Diamond$ and $\boxbar$ in place of $\Box$ 
everywhere. That the theorem goes through is a known theorem of Linnebo in a slightly 
different key.

The resulting theory has some intruiging features. If one restricts one's attention 
to vertical possibility and necessity, one can construct objects that 
satisfy anything you can get in ZFC, for example the existence of uncountable cardinals, 
beth fixed points, and so on. But these large cardinals are always mirages: 
and the mirage can always be revealed 
by appeal to $\circledbslash$-possibility. Anything that might in some possible world 
satisfy the formula saying that it is (say) $\omega_1$ will also not satisfy that 
description in some other possible world. There are no \emph{absolute} uncountables, only 
relative pretenders.

The chief advantage HWP has over SHWP is interpretative power: it can interpret all 
of ZFC, notwithstanding its ultimate commitment to the countability of all things.
This gives HWP the capacity, at least once motivational details are filled in,
to potentially offer a countabilist foundation for all mathematics. This is I think
one of the principle interests of the combination of height and width potentialism.

I have said that the chief advantage of HWP over SHWP is interpretative power. But 
how much of that does it have? Does it have too much? We now turn to addressing 
these issues of consistency in HWP.

It turns out that the methods employed in the previous section generalize fairly 
naturally to cater to systems like HWP and even further extensions. The generalization
involves extend second order arithmetic to include `topological regularity axioms',
and then giving a more nuanced definition of possible world in terms of such an extension.

Let's start with the extension of second order arithmetic involved, 
to include so-called topological regularity axioms. 
What are those?

Well, there are certain `nice' topological properties of sets of reals -- 
things like being Lebesgue measurable, or having the perfect set property -- 
that cannot hold everywhere, at least given the axiom of choice. The issue (repeatedly)
is that the axiom of choice allows you to well-order the reals and then construct
barbaric sets of various kinds, ones that don't have the nice features, by exploiting 
the well-order. 

There's a general feeling in set theory that the `nice' properties should all 
hold of easily definable sets of reals, and that the `nasty' counterexample sets 
should all be pretty complicated to define (in terms of the usual analytical hierarchy).
For example, if $V = L$, then there are easily definable nasty sets of reals, and 
this is generally taken to be a mark against the principle.

Since 
$V = L$ implies there are easily definable but nasty sets, 
it follows that principles asserting 
that easily definable sets are nice must in some cases go beyond SOA (which is 
of course consistent with $V = L$). Principles of this kind 
are what I'm calling \emph{topological 
regularity axioms}. They assert that good-behavior properties hold for 
certain easily definable classes of reals, even when this is not 
provable in SOA (or ZFC). 

The weakest such axiom is something called the
$\Pi_1^1$-Perfect Set Property (PSP), which says that every uncountable set of reals 
definable by a $\Pi_1^1$ formula is either countable or has a perfect subset.\footnote{
    A perfect set is a closed set with no isolated points.
} It turns out that this minimal extension of SOA is enough to secure 
the interpretability of HWP.

\begin{theorem}
    There is a map $\cdot^\exists : \mathcal{L}_1 \to \mathcal{L}_\in$
    that preserves theoremhood from HWP to SOA + $\Pi_1^1$-PSP.
\end{theorem}
The proof uses similar ideas to those employed for SOA and SHWP in the previous 
section, but as I said one has to take a more nuanced definition of `possible world'. 
In particular, it turns out that here our definition of a world will 
need to be `doubly parameterized': that is, we will need \emph{two} free 
variables in $\varphi^\exists$. 

Why?
Let us think about how our possible worlds will have 
to look to get an interpretation going. In order for r-Pow to be true, 
there will have to be worlds $w$ at which the set $x$ 
of all $\circledbar$-possible sets of numbers 
exist. These worlds are worlds where 
any plurality of numbers, in the sense of $w$, forms a set. 
And yet, since we can always force to add subsets to any infinite set, 
there must still be further possible sets of numbers. These sets of numbers 
have numbers as their members. But we said that all pluralities of numbers at 
$w$ formed sets at $w$!

The only way out of this tangle for the HWP proponent is to accept indefinite 
extendability both with respect the singular domain, and with respect the plural 
domain (holding the singular domain fixed). The introduction of new generics 
witnesses, on this view, the extension of the plural domain (over a fixed) singular 
domain. Indefinite extendability `runs in two dimensions', according to the present 
picture. Hence, our possible worlds will be doubly parameterized, with one 
parameter representing the singular domain, and the other parameter representing the
plural. Expansions will then be possible, under the proposed interpretation, in 
both directions. 

Let's now look in more detail 
how this works. We will make use of the following fact, whose proof 
is originally due to Solovay.
\begin{fact}
    Over SOA, the $\Pi_1^1$-PSP is equivalent to $L[r]$ containing only 
    countably many reals for every $r$.
\end{fact}
Thus, in effect, the Fact says that $L[r]$ always falls badly short of containing 
all reals if we have $\Pi_1^1$-PSP. It is thus a strong form of $V \not = L$.

It is fairly easy to see that the fact has the following corollary.
\begin{corollary}
    In  SOA + $\Pi_1^1$-PSP, $L[r]$ is a (class) model of ZFC for every real $r$.
\end{corollary}
I provide an argument sketch (that the corollary follows from the fact) in 
a footnote.\footnote{
    Do.
}

This now gives us the tools we need to implement our parametrized possible worlds 
strategy. We define, in SOA + $\Pi_1^1$-PSP, our `possible worlds' to comprise
a transitive set $t$ (representing the first order domain) and a real number $r$
(representing the second order domain). We require that $t \in L[r]$, 
and will say that plural quantification in a possible world is always 
restricted to subsets of $L[r]$. A possible world $t, r$ may then be extended 
along two paramters. `Vertical' expansion expands the $t$ parameter, staying within 
$L[r]$. `Horizontal' expansion expands the $r$ parameter, and insodoing 
accommodates the needed growth in the plural domain even holding the first 
order domain fixed. $t$ will in general have different subsets in $L[r]$ than 
$L[s]$, and \emph{these} are the first order theorist's interpretation for talk of 
plurals in HWP. 

To implement this formally, 
our mapping $\varphi \mapsto \varphi^\exists$ will therefore 
need to carry $\varphi \in \mathcal{L}_1$ to a formula in $\mathcal{L}_\in$ with 
not just one but two free variables: $\varphi \mapsto \varphi^\exists(t, r)$. It will 
contain clauses like 

\[ (\forall x \varphi)^\exists(t, r) := \forall x \in t \varphi^\exists(t, r) \]
\[ (\forall X \varphi)^\exists(t, r) := \forall x \subseteq t [x \in L[r] \rightarrow \varphi^\exists(t, r)] \]
\[ (\boxbar \varphi)^\exists(t, r) := \forall u \in L[r] [Tran(u) \wedge t \subseteq u \wedge u \in L[r] \rightarrow \varphi^\exists(u, r)]\]
\[ (\Box \varphi)^\exists(t, r) := \forall s \geq r \forall u \in L[s] [Tran(u) \wedge t \subseteq u  \rightarrow \varphi^\exists(u, s)]\]
(In the final expression, $s \geq r$ means that $r$ is constructible from $s$.)

With the full translation in hand 
it is a fairly straightforward matter to prove that every theorem of HWP 
comes good in SOA + $\Pi_1^1$-PSP under the translation: 
$\varphi \mapsto \forall t, r Tran(t) \wedge t \in L[r][ \varphi^\exists(t, r)]$.
(The fact that $L[r]$ models ZFC
in SOA + $\Pi_1^1$-PSP is needed to get the translations of the $\circledbar$ 
axioms to come good.)
In fact, one can show that HWP has $\Pi_1^1$-PSP$^\Diamond$ as a theorem, and 
prove the same kind of tight proof-theoretic equivalence obtains here 
between HWP and SOA + $\Pi_1^1$-PSP as did between SHWP and SOA simpliciter. 
These results can be pushed further: by increasing the strength of the regularity 
assumed in the interpreting theory, 
one can ratchet up the consistency strength of the corresponding modal HWP 
theory, allowing for measurables and beyond to be `vertically' attainable.
(We will return to this observation in more detail below.)

In any case, that concludes the formal arguments 
in favor of the consistency of height and width potentialism.
We have seen, in outline, that there are arguments for the consistency of 
height and width potentialism in both a strong and a weak form, relative to 
extensions of SOA by topological regularity axioms.

Let us 
now turn to the arguments alleging inconsistency between height and width 
potentialism, and to see how they relate to the approaches sketched up to now.

\section{Roberts}

Right from the off it is perhaps worth mentioning 
that Roberts' argument applies to HWP only. This is not a big deal: 
he is assuming that the height potentialist will be committed at least 
to the recovery of ZFC in terms of pure height potentialism, so that 
any commitment to width potentialism will require a bimodal treatment 
of the sort described in the last section. But it is worth mentioning 
that a less grand form of height and width potentialism can still be viable,
in the form of SHWP.

The actual details of Roberts' argument are at present not entirely 
clear to me. His paper is a work in progress, but in the most 
recent version he argues that the addition of an actuality operator 
to a theory related to HWP is incosistent, and that on that basis 
we should think that the viewpoint enshrined in HWP is similarly inconsistent.
In correspondence, I suggested that the use of an actuality operator 
was a red-herring, and that in fact the argument turns (whether or not 
we have an actuality operator in the picture) on whether we are allowed 
comprehension only in the `closed' form cited above, or if instead 
we accept instances of comprehension with free variables. I continue 
to believe that the fundamentals of the argument Roberts presents turn 
on issues surrounding comprehension, and that addition of an actuality 
operator is a needless increase in complexity. So here I will discuss 
what I take to be the simpler version of the argument, rather than 
the last one I saw in a draft from Roberts. In an appendix, I will 
explain why I think the addition of an actuality operator is unnecessary.

Here then is the argument. Let HWP$^+$ be the 
results of allowing all instances of comp, 
\[ \exists X \forall x[Xx \equiv \varphi]\]
where $\varphi$ may include free variables. Then:
\begin{theorem}
    HWP$^+$ is inconsistent.
\end{theorem}
\begin{proof}
    It is easy to see that HWP implies
    \begin{equation}
        \Diamond \exists x[\Diamond (\exists y[ y \subseteq x \wedge y = z]) \wedge \boxbar(\neg \exists y[y = z])]
    \end{equation}
    abbreviate this as $\Diamond \exists x \Psi(x, z)$. By 
    existential instantiation and the comprehension instance
    \begin{equation}\label{exa}
        \exists X \forall w[Xw \equiv \Diamond w \in z]
    \end{equation}
    in parameter $z$ we get:
    \begin{equation}\label{xb}
        \Diamond (Ex \wedge \Psi(x, z) \wedge \exists X \forall w[Xw \equiv \Diamond w \in z])
    \end{equation}
    Then by HP, 
    \begin{equation} 
        \Diamond (Ex \wedge \Psi(x, z) \wedge \circledbar \exists u \forall w[w \in u \equiv \Diamond w \in z])
    \end{equation}
    one can then instantiate on $u$ and prove $u = z$ by 
    (our strong modal form of) extensionality.

    But this is a contradiction.

\end{proof}

Naturally, given this argument, we should expect the proof of consistency implicit in the 
previous section to break down with strong comprehension. And this is precisely 
what happens in the interpretation offered for HWP in SOA $+ \Pi_1^1$-PSP.
Consider for example the following simple case. Let $w$ be the world that has $\omega$
for its first order domain and $P(\omega)^L$ for its plural domain. 
The interpretation for closed plural comprehension at this world amounts to the claim 
that for any $y$ and $Y$ \emph{in $L$}, the subset of $\omega$ for which 
$\varphi(n, y, Y)$ holds is again in $P(\omega)^L$, which is obviously true. But 
if we remove the restriction to closed formulas, we are saying that for any $\varphi$, $y$ and $Y$ 
we choose (whether in $L$ or not), the set of $n$ with $\varphi(n, y, Y)$ is in $L$. 
But this is provably false in SOA + $\Pi_1^1$-PSP.

The question about the consistency of HWP that arises from this argument concerns, then,
the status of the `strong' open version of plural comprehension. Formally speaking 
the path toward saving HWP is clear: accept only the closed instances of comprehension.

Roberts' claim is that the open instances are mandatory:
any reasonable plural logic will contain them. But I see no good reason 
for thinking this is so. 

For one thing, it is generally assumed that formulas with free 
variables at best play an auxiliary role in such formalizations, since 
under interpretation they don't say anything at all. They are analogous to 
assertions with hanging pronouns, like in `it's a good thing'. To rest a philosophical 
case on such formulas being mandatory would therefore be peculiar.

For another, I would argue that the intuitive validity 
of the instances of comprehension that are used in 
the cited argument is legitimately (and independently) questionable, 
whether or not one is a potentialist. In particular, 
those of a Stalnakerian actualist bent in modal metaphysics 
are likely to find the open schema doubtful.

Let me explain. Stalnakerian actualism involves contingentism about 
(among other things) \emph{properties}. For example, 
in the event that I should not exist, the property of not being 
identical to me would also fail to exist according to Stalnaker. Of course, 
given that I exist, it makes sense to look upon some 
other circumstances in which I don't exist, and truly affirm 
of various things that they are not identical to me. 
But this does not mean that the world in question would 
be one in which various things had the property of not being 
identical to me, since were those circumstances to be actual 
there would be no such property. 

Whatever the merits of this view in general -- I myself 
happen to be sympathetic to it -- it is clearly coherent and 
a reasonable idea to consider in modal metaphysics. At its 
core, from a formal point of view, is precisely a denial of 
arbitrary instances of comprehension for properties. It is 
not true, on this picture, that arbitrary worlds will satisfy comprehension for 
properties on formulas like `being identical to $x$', though 
they will have the closed analogue (for every $x$ there is a property 
of being identical to $x$.) Similar points go for `higher order' instances 
of comprehension, like `being identical to the property $P$'.

For a Stalnakerian, the problem with open instances of property comprehension
is that they allow worlds illict access to things in \emph{other} 
worlds in the model.
In our formal model, we have a bunch of worlds with separate domains,
and we can name all the things in the domain of any model and ask 
questions about whether one thing that exists according to one world 
exists according to another. But these questions are only askable with 
our external, model-theoretic perspective. 
When $z$ gets assigned something that doesn't exist at $w$, 
inhabitants of $w$ do not have the resources to refer to $z$, 
and there in a certain sense simply are no comprehension instances 
involving $z$ at $w$.
Such formulas, 
under interpretation in the model,
just don't correspond to real propositions that exist at the 
worlds at which truth is to be evaluated.
There is no thought there to be thought there.

The view transposes straightforwardly to one that 
denies the relevant instances of comprehension for plurals.
Presumably, on this Stalnakerian view, in circumstances where 
I do not exist there are no such things as my molecules, 
whether or not my molecules as we now recognize them
should in fact exist at the world in question.  Similarly, 
in circumstances where a generic filter
for a partial order does not exist there are no such things 
as its elements, whether or not the elements we will eventually 
recognize it to have in fact exist at the world in question.

Thus the Stalnakerian actualist 
has independent grounds 
for restricting the allowable instances of comprehension to be closed. This amounts 
to only allowing parameters that represent actual things on comprehension. But from 
a Stalnakerian point of view this is no restriction at all. The other instances 
do not even correspond to genuine thoughts.

This seems to me to be a clear enough rationale for holding only closed comprehension 
to be valid. But the alleged inconsistency argument turns on the open instances.
I conclude it does not go through.

All told then I do not think Roberts' argument shows HWP is inconsistent.
But I should say that the argument does do an excellent job 
at bringing out the contours of the philosophical commitments that go along 
with the view, and in this regard I think it is still a significant contribution.

There is a related argument in the vicinity here that is also worth some brief 
discussion. This says that HWP is inconsistent because it violates the valid 
principle of plural logic
\begin{description}
    \item[BF] $\Box \forall Y (\Diamond \exists X \sqsubseteq Y [X = Z] \supset \exists X \sqsubseteq Y[X = Z])$
\end{description}
where here $X \sqsubseteq Y$ abbreviates $\forall x[Xx \supset Yx]$. 
This says that given pluralities cannot gain sub-pluralities, and is violated 
in HWP. Moreover, BF is motivated on standard conceptions of plural 
logic according to which a plurality is `nothing over and above' 
the things it comprises. So, one might conclude, HWP is inconsistent after all.

I agree that BF is well-motivated on that conception but do not believe 
the conception is sacrosanct. Our conception of plurals -- of `some things' -- 
is one bequeathed us from informal thought. Precisely what logical principles 
they should obey, and how they are to be conceptualized, is not rigidly 
determined \emph{a priori}, but something that is to a great extent up to 
us to decide how to fill in the details. Indeed there are in all likelihood 
a number of related notions that can serve different roles in philosophical 
enquiry. The task here is to find one that suits HWP, and the one that does 
will necessarily not satisfy BF, since given any definite range of pluralities 
over a given domain one can find new ones by forcing. If the reader is really unhappy 
with this being true for `pluralities', I'm happy to drop the term: call them 
strongly extensional properties if you like, it doesn't matter, so long as they 
can plausibly be held to have the properties required of them by the axioms 
of HWP.

This point is one of general significance.
HWP uses a cluster of ideas: modality, set membership, quantification, plurality, 
and so on. None of them is an absolute fixed point, and there are basically no 
conceptual fixed points within them. It is a matter of finding the right configuration 
of these things to do justice to the guiding picture. That is why, in my view,
it is highly unlikely that any serious inconsistency argument for the combination of
height and width potentialism will be forthcoming: 
pretty much any derived inconsistency  in a formal system would as naturally 
be taken to show 
the formalization was not up to the task. This 
generally seems to be the right response to these sorts of inconsistency proofs.\footnote{
    Compare e.g. Russell's paradox and the concept of set, or Berkeley's 
critique of the notion of the infinitesimal.}

\section{Brauer}
Brauer's argument is quite closely related to Roberts', and in a sense the 
same key logical observations are at work in each. But Brauer's argument brings 
out some interesting issues distinctive to the foundations of set theory 
that are not really brought up by Roberts, and for that reason 
warrants seprate discussion.

The core argument is simple.
\begin{description}
    \item[P1] Height potentialism, as motivated by the iterative concept of set,
    sanctions only convergent extensions of the universe: if it could be that 
    the set $x$ existed, and it could be that some other set $y$ existed, 
    then it could be that $x$ and $y$ 
    both exist together.  
    \item[P2] Width potentialism, as motivated by the idea of forcing potentilism,
    requires allowing for divergent extensions of the universe: 
    for there are (or at least could be) partial orders 
    $P$ and possible generic filters $G_1$ and $G_2$ such that $G_1$ and $G_2$ 
    could not exist together.
    \item[C] Height and width potentialism (so understood) are inconsistent.
\end{description}
I'm not completely sold on either premise. In the case of P1,
Brauer gives an argument which he glosses informally as follows:
\begin{quote}
    The basic idea of the argument is straightforward: 
    if you have a universe $V_0$, you can get different extensions 
    $V_1$ and $V_2$ when $V_1$ 
    is formed by bringing together some collections $X$ into sets and 
    $V_2$ is formed by bringing some other collections $Y$ into sets; 
    but nothing in the process of forming sets out of the 
    $X$ collections would then prevent you from 
    later forming sets out of the $Y$ collections.
\end{quote}
So, we are to conclude, that according to height potentialism under the guise of
the iterative concept of set, only convergent extensions of the universe are possible.

My doubts about this form of argument concern whether it is OK to conclude 
that quite generally the iterative concept of set is incompatible with 
divergent extensions of the universe. After all, set theory is not everything, 
and although there may be nothing in the simple process of forming sets 
from given things that prevents 
re-convergence, nevertheless other things may do so. 

For example, suppose I have a handle, a blade, some permanently 
binding superglue, and a universe of sets $V$. I can then form the knife $K_1$
and form its singleton getting a universe of sets $V_1$. Or I can form 
the knife $K_2$ and form \emph{its} singleton getting $V_2$. But these 
can never be unified. 

Or, for a more pointed example (in the spirit of the `second approach' to width 
potentialism discussed by Brauer on p7), suppose I have two machines. 
One takes in some things and gives back a set with them as its elements.
The other takes in a 
partial order, some dense sets in it, and gives back a plurality that 
intersects all the given dense sets. Then, given suitable background assumptions 
(including, but not by any means restricted to P2),
we will be in a structurally analogous situation to the blade-handle 
case: we might find it is possible to produce a set whose elements are $G_1$,
and also possible to find a set whose elements are $G_2$,
without these addmiting a unification.
But this is no different from the previous case, which is (hopefully?) completely 
unproblematic.

But discussion of P1 is really a red herring, since at least so far as the 
brands of height and width potentialism here go P2 is false:
one can argue in (S)HWP that if it is possible that some generic 
$g_1$ exists, and it is possible that some generic $g_2$ exists, then it is 
possible that they exist together. One can see this most easily by using 
the tight form of equivalences offered with extensions of SOA. 
Such theories prove that
any two generics for a partial order are both contained in a common transitive 
set, for example a sufficiently large fragment of $L[r]$ with $r$ witnessing 
the countability of all the relevant parties.
Thus the modal logic of forcing, in this 
setting, \emph{is} convergent. 

Brauer's argument on the other hand is based on the familiar result from ZFC saying 
that, in any given countable transitive model for ZFC, one can find a partial 
order and a pair of generics for it that can't exist together in a model of ZFC, 
because they witness the countability of the model in question (Woodin). This
is clearly correct, and is indeed a fact that the proponent of HWP can 
recognize: they can see for example that it is vertically possible that there
are models for (fragments of) ZFC, and partial orders they contain, and such that no 
possible model for the same theory contains both generics. But from the point of 
view of HWP, this is just because you are ignoring some universes, namely those 
wherein everything under discussion is in fact countable (and so ZFC is false).

The point is that HWP is associated with a radically different coneption of 
the mathematical universe than that associated with ZFC. Now I must concede 
that there is a sense in which the iterative concept of set is tied very closely 
to ZFC and to certain forms of height potentialism: and if one thinks that 
these links are essential to height potentialism then of course it is right 
to say that height potentialism is inconsistent with width potentialism.  
That is, if you think height potentialism requires commitment 
to the potential hierarchy being stratified into $V_\alpha$s that 
satisfy a modal analogue of ZFC, then trivially we will find that 
width potentialism is ruled out.
(This barely requires argument.) But in any case
I see no reason to think such a thing. The core idea of height potentialism is 
that the universe can always be extended upward, not that it can't be extended outward
too. If fleshing out that position means temporarily giving up on ZFC, so be it.
(Though as we have seen ZFC is readily interpretable in HWP in something resembling 
`innner models'.)

I hope it is clear that there is 
at least a coherent 
 position that incorporates a lot of the traditional ideas in the 
iterative conception, understood in height potentialist terms, but that amalgamates 
width potentialism as well. 





\end{document}